A Domain-Specific Optimizing Compiler\\for Solving Partial Differential Equations

### 2014-02-18 paper/bare_conf.tex
Add pictures of normalized runtimes
Start writing Background section

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

We focus on Finite Element Local Assembly (``assembly'', in the following), a fundamental step of FEM that covers an important fraction of the overall computation run-time, typically in the range 30$\%$-60$\%$. During the assembly phase, the solution of the PDE by executing a suitable kernel over all elements in the discretized domain. A kernel's working set is usually small enough to fit the L1 cache; it might need L2 cache when (very) high-order methods are employed to improve the accuracy of the solution. However, we do not consider the latter case. An assembly kernel is characterized by the presence of an affine, usually non-perfect loop nest, where individual loops are rather small (the trip count rarely exceeds 100, with a minimum value of 3, depending on the order of the method). With such small kernels, our work focuses on aspects like minimization of floating-point operations, register allocation and instruction-level parallelism, especially in the form of SIMD vectorisation.

@@

\section{Background}

%Here we discuss about the computational characteristics of computational science kernels. Briefly cite op2 kernels. Emphasis on Finite Element Assembly. Generalization and formalization of a Finite Element Assembly kernel using quadrature representation.

@@

\section{Performance Evaluation}
%Performance evaluation. We show all of the results got so far and we explain them. We demonstrate the claim that a compiler-driven approach is necessary to obtain maximum performance. Also, our results prove that current vendor compilers still have notable limitations, despite optimizing loop nests that are 1) affine, 2) have no stencil, and 3) ?

>>>>

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

@@

We focus on Finite Element Local Assembly (``assembly'', in the following), a fundamental step of FEM that covers an important fraction of the overall computation run-time, typically in the range 30$\%$-60$\%$. During the assembly phase, the solution of the PDE is approximated by executing a suitable kernel over all elements in the discretized domain. A kernel's working set is usually small enough to fit the L1 cache; it might need L2 cache when (very) high-order methods are employed to improve the accuracy of the solution. However, we do not consider the latter case. An assembly kernel is characterized by the presence of an affine, usually non-perfect loop nest, where individual loops are rather small (the trip count rarely exceeds 60, with a minimum value of 3, depending on the order of the method). With such small kernels, our work focuses on aspects like minimization of floating-point operations, register allocation and instruction-level parallelism, especially in the form of SIMD vectorisation.

@@

\section{Background}
%Here we discuss about the computational characteristics of computational science kernels. Briefly cite op2 kernels. Emphasis on Finite Element Assembly. Generalization and formalization of a Finite Element Assembly kernel using quadrature representation.

Local assembly consists of evaluating so called element stiffness matrix and vector; in this work, we focus on computation of matrices, which is the costly part of the process. A stiffness matrix can be intuitively thought as an approximated representation of the PDE solution in a specific cell of the discretized domain. Quadrature algorithms (based on numerical integration), which follow on from the mathematical description of a stiffness matrix, are widely-used in scientific codes to implement assembly~\cite{quadrature1, quadrature 2}. The FFC compiler used by Firedrake and Fenics can generate assembly kernels using optimized quadrature schemes. It is worth noticing, however, that problems and considerations similar to that exposed below can be found also in non-FFC assembly codes.

The complexity of FFC-generated kernels depends on the mathematical problem being solved. In simpler cases, the loop nest (always characterized by affine memory accesses) is perfect, it has short trip counts (in the range 3-15), and the computation reduces to a summation of a few, small outer-product vector-operations. An example is provided in Figure~\ref{helmholtz-code}, which shows an assembly kernel for a Helmholtz problem, using Lagrange basis functions on 2D elements with polynomial order $p=1$. In other situations, for instance when solving a non-linear problem like Burgers (see Figure~\ref{burgers-code}), the number of arrays involved in the computation of the stiffness matrix can be much larger (14 in the case of Burgers using $p=1$ Lagrange tetrahedrons). Also, constants evaluated in outer loops, acting as scaling factors of arrays, may be required; trip counts can be larger (the increment is proportional to the order of the method); arrays may be block-sparse. Despite such a big space of possible assembly implementations, it is still possible to identify a domain-specific structure common to all kernels, which can be turned into effective code transformations and simd-vectorization.

%

@@

\section{Performance Evaluation}
%Performance evaluation. We show all of the results got so far and we explain them. We demonstrate the claim that a compiler-driven approach is necessary to obtain maximum performance. Also, our results prove that current vendor compilers still have notable limitations, despite optimizing loop nests that are 1) affine, 2) have no stencil, and 3) ?
\begin{figure}[h]
\centerline{\includegraphics[scale=0.7]{Pictures/helmholtz-normalized.pdf}
\label{fig_first_case}}
\caption{Helmholtz}
\end{figure}

\begin{figure}[h]
\centerline{\includegraphics[scale=0.7]{Pictures/advdiff-normalized.pdf}
\label{fig_first_case}}
\caption{Advection-Diffusion}
\end{figure}

\begin{figure}[h]
\centerline{\includegraphics[scale=0.7]{Pictures/burgers-normalized.pdf}
\label{fig_first_case}}
\caption{Burgers}
\end{figure}

%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{Pictures/helmholtz-normalized.pdf}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{Pictures/advdiff-normalized.pdf}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
